---
title: "Homework 6"
author: "46-923, Fall 2017"
date: "Due Thursday, December 14 at 3:00 PM"
output: pdf_document
---

\large

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(quantmod)
library(tseries)
```

You should complete this homework solely within R Markdown.
You do not need to show any derivations that were required
to complete Question 1, just implement the analysis in R.

\vspace{.2in}

__Question 1__ 

Assume that $X_1, X_2, \ldots, X_n$ are iid from the Poisson$(\lambda)$
distribution.

Assume you want to test $H_0\!: \lambda = 4.3$ versus $H_1\!: \lambda < 4.3$
using the Wald test.
Calculate the p-value when the data are those found via 
```{r,eval=FALSE}
x = read.table("http://www.stat.cmu.edu/~cschafer/MSCF/PoisHypTest.txt")
```


\vspace{.2in}

__Question 2__

The __Augmented Dickey-Fuller (ADF) Test__ is a test utilized
in time series analysis in order to assess stationarity.
There is a function `adf.test()` in the package `tseries`
which implements this test.
Take a look at `help(adf.test)` for some details.

a. If our objective is to see if there is strong evidence
that a time series is stationary, how should the
argument `alternative` be set when using `adf.test()`?

This is basically asking about the null / alternative hypotheses. The null hypothesis, $H_0$, is "the time-series is stationary"; the alternative hypothesis, $H_1$, is the "the time series is not stationary". Hence, the argument `alternative` should be set to `explosive`. 

b. Consider the following R commands. This will read in the
stock data discussed in lecture, and then run the ADF test
on the first stock in the sample. Use this as a starting point
to write code to loop over all 1000 stocks and get the p-values
for each test. How many of the tests have p-values less than 0.05?

```{r,eval=FALSE}
stocksample = read.table("stocksample.txt", header=T,
                          sep="\t", comment.char="")

adf.test(as.numeric(log(stocksample[1,5:34])), alternative="s")
```

c. Run the p-values found in part (b) through the Benjamini-Hochberg
procedure described in lecture. Are any of the series found to be
stationary using this approach (again using $\alpha=0.05$)? Comment on the reason(s) for any differences found.

