---
title: "Homework 2"
author: "46-923, Fall 2017"
date: "Due Thursday, November 9 at 3:00 PM"
output: pdf_document
---

\large

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(quantmod)
```

You should submit the Rmd and pdf file for Question 1.
You should also submit a pdf file with your responses to
Questions 2 through 4. There is nothing wrong with handwritten
solutions; I am not asking you to learn Latex to complete
the homework.

__Please do not submit photos of your homework.__ Scanners
are available for your use.

\vspace{.2in}

__Question 1:__

Run a __both__ K-means and hierarchical clustering algorithm on the
yield curve shift data that was considered back when PCA was
introduced. Discuss anything interesting that you find. You are free
to make decisions regarding settings to the algorithms as you see fit.

```{r, warning=FALSE}
## Scrape the data
library(ggplot2)
library(xml2)
library(rvest)
library(diffusionMap)
library(reshape2)

fullYCweb =read_html("https://goo.gl/j97141")
tvdnodes =html_nodes(fullYCweb, ".text_view_data")
tableelements =html_text(tvdnodes)

#Change N/A in NA
tableelements[grep("N/A", tableelements)] = NA
YCdata =matrix(tableelements, ncol=12,byrow=TRUE)
YCdata =data.frame(YCdata, stringsAsFactors=FALSE)
names(YCdata) =c("Date","1mo","3mo","6mo","1yr",
                 "2yr","3yr","5yr","7yr","10yr",
                 "20yr","30yr")
YCdata$Date =as.Date(YCdata$Date,format="%m/%d/%y")
YCdata[,2:12] =apply(YCdata[,2:12],2,as.numeric)
YCrates = YCdata[YCdata$Date > "2010-01-01", -1]
YCrates = YCrates[-which(apply(YCrates,1,sum) == 0),]
YCshifts = apply(YCrates,2,diff)

## K-Means against dimension reduction
km_out = kmeans(YCshifts[complete.cases(YCshifts),], centers=5, nstart=10)
YCshifts_cpy = YCshifts[complete.cases(YCshifts),]
YCshifts$km_clust = factor(km_out$cluster)
YCshifts = data.frame(YCshifts)
dist_Map = dist(YCshifts_cpy[complete.cases(YCshifts_cpy),])
diff_map = diffuse(dist_Map, eps.val=50, t=10)
YCshifts$diff_1 = diff_map$X[,1]
YCshifts$diff_2 = diff_map$X[,2]
YCshifts$clust = km_out$cluster

## convert to df for ggplot
YCshifts = data.frame(YCshifts)

plt = ggplot(YCshifts, aes(x=diff_1, y=diff_2, colour=km_clust)) +
  geom_point() +
  labs(x="First Diffusion Coord", y="Second Diffusion Coord", color="Cluster")

plt

## Hierarchical
hcout = hclust(dist(YCshifts_cpy[1:50,]), method="complete")
plot(hcout, cex=0.35, xlab="")


```


