---
title: "Homework 5"
author: "Jordan Giebas"
date: "Due Thursday, November 30 at 3:00 PM"
output: pdf_document
---

\large

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(quantmod)
```

__Please do not submit photos of your homework.__ Scanners
are available for your use.

\vspace{.2in}

__Question 1__ 

```{r}
options("getSymbols.yahoo.warning"=FALSE)
options("getSymbols.warning4.0"=FALSE)
source("http://www.stat.cmu.edu/~cschafer/MSCF/ModelSelectionExample.txt")
library(fGarch)
library(ggplot2)
library(ggfortify)

Kellogg = getSymbols("K", from="2010-1-1", to="2016-12-31", auto.assign=F)
lwrk = data.frame(weeklyReturn(Ad(Kellogg), type="log"))

# Generalized Error Distribution
print("Generalized Error Distribution trial")
ged_out = FitGED(lwrk$weekly.returns)
print(ged_out$mle)

# Non-standard T-distribution
print("Non-standard T-distribution trial")
nstd_out = FitNonStdT(lwrk$weekly.returns)
print(nstd_out$mle)

# Skewed T-distribution (no shift)
print("Skewed T-distribution (shifted, optimized) trial")
skt_out = FitSkewT(lwrk$weekly.returns)
print(skt_out$mle)

# Skewed T-distribution (shifted, optimized)
print("Skewed T-distribution (shifted, optimized) trial")
sktShiftOpt_out = FitSkewT(lwrk$weekly.returns, allowshift=T, control=list(maxit=1000))
print(round(sktShiftOpt_out$mle,6))

# Normal Distribution
print("Normal Distribution trial")
n = length(lwrk$weekly.returns)
norm_out = c(mean(lwrk$weekly.returns), sqrt(var(lwrk$weekly.returns)*(n-1)/n))
print(norm_out)

### For each trial, compute LLH

# Generalized Error Distribution
print("Generalized Error Distribution plot/llh")
print(sum(dged(lwrk$weekly.returns, ged_out$mle[1], ged_out$mle[2], ged_out$mle[3], log=TRUE)))

# Non-standard T-distribution
print("Non-standard T-distribution llh")
print(sum(dstd(lwrk$weekly.returns, nstd_out$mle[1], nstd_out$mle[2], nstd_out$mle[3], log=TRUE)))

# Skewed T-distribution (no shift)
print("Skewed T-distribution (no shift) llh")
print(sum(dSkewT(lwrk$weekly.returns, skt_out$mle[1], skt_out$mle[2], skt_out$mle[3], skt_out$mle[4], log=TRUE)))

# Skewed T-distribution (shifted, optimized)
print("Skewed T-distribution (shifted, optimized) llh")
print(sum(dSkewT(lwrk$weekly.returns, sktShiftOpt_out$mle[1], sktShiftOpt_out$mle[2], sktShiftOpt_out$mle[3], sktShiftOpt_out$mle[4], sktShiftOpt_out$mle[5], log=TRUE)))

# Normal Distribution
print("Normal Distribution llh")
print(sum(dnorm(lwrk$weekly.returns, norm_out[1], norm_out[2], log=TRUE)))


### Constructing plots, code taken from Model_Selection.txt from Canvas
ggplot(lwrk,aes(x=weekly.returns)) +
  geom_density(bw="SJ",aes(color="kde")) +
  stat_function(fun=dnorm, aes(color="normal"),
                args=list(mean=norm_out[1],sd=norm_out[2])) +
  stat_function(fun=dged, aes(color="ged"),
                args=list(mean=ged_out$mle[1],sd=ged_out$mle[2],nu=ged_out$mle[3])) +
  stat_function(fun=dstd, aes(color="nonstdT"),
                args=list(mean=nstd_out$mle[1],sd=nstd_out$mle[2],nu=nstd_out$mle[3])) +
  stat_function(fun=dSkewT, aes(color="skewT"),
                args=list(k=skt_out$mle[1],n=skt_out$mle[2],lambda=skt_out$mle[3],sigma2=skt_out$mle[4])) +
  stat_function(fun=dSkewT, aes(color="skewTshift"),
                args=list(k=sktShiftOpt_out$mle[1],n=sktShiftOpt_out$mle[2],lambda=sktShiftOpt_out$mle[3],sigma2=sktShiftOpt_out$mle[4],shift=sktShiftOpt_out$mle[5])) +
  scale_color_manual(name="",
                     values=c("kde"="black","normal"="blue",
                              "ged"="yellow","nonstdT"="green",
                              "skewT"="red","skewTshift"="brown")) +
  labs(x="Log Daily Return",title="Data for Kellogg's (K)",subtitle="January 2010 through December 2016") +
  scale_y_log10()


## It seems that the Non-Standard T-Distribution has the maximal LLH... doesn't necessarily mean it's the best though.

### Determine Minimal AIC

print("Finding minimal AIC")

# Parameter Vector
p_vec = c(3,3,4,5,2)

# LLH Vector for each
ged_llh = sum(dged(lwrk$weekly.returns, ged_out$mle[1], ged_out$mle[2], ged_out$mle[3], log=TRUE))
nstd_llh = sum(dstd(lwrk$weekly.returns, nstd_out$mle[1], nstd_out$mle[2], nstd_out$mle[3], log=TRUE))
skwt_llh = sum(dSkewT(lwrk$weekly.returns, skt_out$mle[1], skt_out$mle[2], skt_out$mle[3], skt_out$mle[4], log=TRUE))
skwtShiftOpt_llh = sum(dSkewT(lwrk$weekly.returns, sktShiftOpt_out$mle[1], sktShiftOpt_out$mle[2], sktShiftOpt_out$mle[3], sktShiftOpt_out$mle[4], log=TRUE))
norm_llh = sum(dnorm(lwrk$weekly.returns, norm_out[1], norm_out[2], log=TRUE))

llh_vec = c(ged_llh, nstd_llh, skwt_llh, skwtShiftOpt_llh, norm_llh)

aic_func = function(p,llh) {
  
  return ( -2*llh + 2*p ) 
  
}

aic_vec = c()
for ( j in 1:length(p_vec) ) {
  aic_vec[j] = aic_func(p_vec[j], llh_vec[j])
}

print("aic_vec:")
print(aic_vec)
print("Min of aic_vec:")
print(min(aic_vec))
print("Max of llh_vec:")
print(max(llh_vec))


```

By AIC, the model selected is the Non-Standard T-distribution since it is the minimal AIC out of the models compared. It has an AIC value of -1850.368. Additionally, the model that yielded the maximal loglikelihood was the Non-Standard T-distribution as well. Therefore, it seems that the Non-Standard T-distribution would be the best model for modelling log returns, perhaps the usage of the Normal Distribution is just because of its ease. 

__Question 2__ 

Scanned in jgiebas_HW5.pdf

__Question 3__ 

Scanned in jgiebas_HW5.pdf











